A Realistic Guide to Making Data Available Alongside Code to Improve Reproducibility

### 2019-06-11 paper.Rmd
add slugs to headings so they can be referenced as sections in text. Fixes minor reference bug and moves something to morgue

# Introduction

Data are a fundamental currency upon which scientific discoveries are made. Without access to good data, it becomes extremely difficult, if not impossible, to advance science. Yet, a large majority of data on which published research papers are based rarely see the light of day and are only visible to the original authors [@Rowhani-Farid2016; @Stodden2018].  Sharing data sets upon publication of a research paper has benefits to individual researchers, often through increased visibility of the work [Popkin2019; @Kirk2019]. But there is also a component of benefit to the broader scientific community. This primarily comes in the form of potential for reuse in other contexts along use for training and teaching (@McKiernan2016). Assuming that the data have no privacy concerns (e.g., human subjects, locations of critically endangered species), or that the act of sharing doesn't put the authors at a competitive disadvantage, sharing data will always have a net positive benefit. First and foremost, sharing data along with other artifacts can make it easier for a reader to independently reproduce the results, thereby increasing transparency and trust in the work. The act of easy data sharing can also improves model training, as many different models can be tried and tested on latest data sources, closing the loop on research and application of statistical techniques. Existing data sets can be combined or linked with new or existing data, fostering the development and synthesis of new ideas and research areas. The biggest of these benefits is the overall increase in reproducibility.

@@

# Challenges in documenting your dataset

@@

# Releasing data in R

@@

# Data Sharing in Julia and Python

@@

# Dealing with medium to large data

@@

# Publishing and repositories

@@

# What belongs in the minimal structure for researchers

@@

## README: A Human readable description of the data

@@

## Codebook: Human readable dictionary of data contents

@@

## License: How to use and share the data

@@

## Citation: How you want your data to be cited

@@

## Machine readable metadata

@@

## Raw data: The original/first data provided

@@

## Scripts: To clean raw data ready for analysis

@@

## Analysis ready data: Final data used in analysis

@@

# Tooling for packaging data

@@

# Example datasets

@@

## Dataset 1: Forest Census Plot on Barro Colorado Island

@@

## Dataset 2: Sensor data

@@

The SDSS provides data at different levels of complexity for different degrees of understanding. It is a huge project involving many people, and a staggering effort to prepare. Despite this, it is still very easy to understand. This is a further reflection on the idea that high effort  can create highly understandable data, (see Figure \@ref(fig:fig-effort-understanding)). The impact of this data is high, having changed the way we understand the past, present and future of the universe itself. This impact is surely due to the care and thought put into making the data accessible and public. It is a statement of what is possible with good data sharing practices.
## Dataset 3: Most other datasets

@@

## Summary

@@

# Ten simple rules for publishing data

@@

# Conclusions

@@

# Morgue

One language agnostic solution is to use the Open Knowledge Foundations Frictionless data spec, implemented in both R and Python.

>>>>

# Introduction {#intro}

Data are a fundamental currency upon which scientific discoveries are made. Without access to good data, it becomes extremely difficult, if not impossible, to advance science. Yet, a large majority of data on which published research papers are based rarely see the light of day and are only visible to the original authors [@Rowhani-Farid2016; @Stodden2018].  Sharing data sets upon publication of a research paper has benefits to individual researchers, often through increased visibility of the work [@ Popkin2019; @Kirk2019]. But there is also a component of benefit to the broader scientific community. This primarily comes in the form of potential for reuse in other contexts along use for training and teaching (@McKiernan2016). Assuming that the data have no privacy concerns (e.g., human subjects, locations of critically endangered species), or that the act of sharing doesn't put the authors at a competitive disadvantage, sharing data will always have a net positive benefit. First and foremost, sharing data along with other artifacts can make it easier for a reader to independently reproduce the results, thereby increasing transparency and trust in the work. The act of easy data sharing can also improves model training, as many different models can be tried and tested on latest data sources, closing the loop on research and application of statistical techniques. Existing data sets can be combined or linked with new or existing data, fostering the development and synthesis of new ideas and research areas. The biggest of these benefits is the overall increase in reproducibility.

@@

# Challenges in documenting your dataset {#doco-challenge}

@@

# Releasing data in R {#data-release-r}

@@

# Data Sharing in Julia and Python {#julia-python}

@@

# Dealing with medium to large data {#med-large}

@@

# Publishing and repositories {#publish-repos}

@@

# What belongs in the minimal structure for researchers {#minimal}

@@

## README: A Human readable description of the data {#readme}

@@

## Codebook: Human readable dictionary of data contents {#codebook}

@@

## License: How to use and share the data {#license}

@@

## Citation: How you want your data to be cited {#cite}

@@

## Machine readable metadata {#machine-metadata}

@@

## Raw data: The original or first data provided {#raw-data}

@@

## Scripts: To clean raw data ready for analysis {#scripts}

@@

## Analysis ready data: Final data used in analysis {#final-data}

@@

# Tooling for packaging data {#tooling}

@@

# Example datasets {#example-data}

@@

## Dataset 1: Forest Census Plot on Barro Colorado Island {#ex-data-1}

@@

## Dataset 2: Sensor data {#sdss-data}

@@

The SDSS provides data at different levels of complexity for different degrees of understanding. It is a huge project involving many people, and a staggering effort to prepare. Despite this, it is still very easy to understand. This is a further reflection on the idea that high effort  can create highly understandable data, (see Figure \@ref(fig:fig-effort-understanding)). The impact of this data is high, having changed the way we understand the past, present and future of the universe itself. This impact is surely due to the care and thought put into making the data accessible and public. It is a statement of what is possible with good data sharing practices.

## Dataset 3: Most other datasets {#ex-data-3}

@@

## Summary {#ex-data-summary}

@@

# Ten simple rules for publishing data {#10-rules}

@@

# Conclusions {#conclusions}

@@
